---
title: "Experimental modalities - Gorilla versus Matlab"
author: "Minah Chang"
date: "1 March, 2022"
output: html_document
---

### Data preparation

From the data, "distractor trials" are discarded and not considered for further analyses in both experimental modalities.

* All Distractor trials were removed (4186 values)

```{r setup, include=FALSE}
# Load and install (if missing) the recquired packages
list.of.packages <- c("tidyverse", 
                      "knitr", 
                      "reshape2", 
                      "lme4",
                      "nlme",
                      "car")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(list.of.packages, library, character.only = TRUE)

# set script folder as working dir
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# set to true to show the code chunk
knitr::opts_chunk$set(echo = TRUE)
```

### Load processed and merged data

The header column is organized as follows:

* 1. pseudoid
* 2. group (EB or SC)
* 3. session (matlab or gorilla)
* 4. category (integer 1, 3, 4, 6-9)
    + category 1: size
    + category 3: brightness
    + category 4: loudness
    + category 6: velocity
    + category 7: movement
    + category 8: space
    + category 9: distance
* 5. wordcode
* 6. language
* 7. stimulus
* 8. meaning
* 9. value (good or goodop)
* 10. response
* 11. grade (1 if correct, 0 if incorrect)

*Native accuracy: given that some words in specific native languages groups can "sound" with the opposite meaning as presented in Tzeng (2017), we recompute the accuracy for those stimuli that in our prestudy were mapped systematically to the opposite meaning. For these subset of stimuli, we consider the response correct if it does not match the original meaning and incorrect if it does.

```{r}

allData<-read.csv("data/allData.csv", header = TRUE)

```

Check that all the participants have the same number of responses per category

```{r}

allData %>% 
  group_by(pseudoid, session, group, category) %>% 
  count() %>% 
  pivot_wider(names_from = category, values_from = n) %>% 
  print(n = 46)

```

### Analysis

### Compute the raw and native accuracy for Gorilla and Matlab.

```{r}

allData$rawAccuracy <- as.integer(allData$response == allData$meaning)

allData %>%
  group_by(session) %>%
  summarize(rawAccuracy = mean(rawAccuracy), 
            .groups = 'keep')

allData$nativeAccuracy <- ifelse(allData$value == 'goodop', 
                                 abs(allData$rawAccuracy - 1), 
                                 allData$rawAccuracy + 0)

allData %>%
  group_by(session) %>%
  summarize(native_acc = mean(nativeAccuracy), 
            .groups = 'keep')

allData %>%
  group_by(session) %>%
  summarize(raw_acc = mean(rawAccuracy), 
            native_acc = mean(nativeAccuracy), 
            .groups = 'keep')

```

### Accuracy per group and category

Category 1: size
Category 3: brightness
Category 4: loudness
Category 6: velocity
Category 7: movement
Category 8: space
Category 9: distance

### Compute the raw and native accuracy for Gorilla and Matlab per group (blind and sighted).
```{r}

categoryAccuracy_wide <- allData %>%
  group_by(group, session) %>%
  summarize(rawAccuracy = mean(rawAccuracy), nativeAccuracy = mean(nativeAccuracy), 
            .groups = 'keep') %>% 
  pivot_wider(names_from = session, values_from = c(rawAccuracy, nativeAccuracy))

print(categoryAccuracy_wide)

```

## Plot of the native accuracy per session for both blind and sighted groups.

```{r}

categoryAccuracy_tidy <- allData %>%
  group_by(session, group) %>%
  summarize(rawAccuracy_mean = mean(rawAccuracy), nativeAccuracy_mean = mean(nativeAccuracy), 
            rawAccuracy_sd = sd(rawAccuracy), nativeAccuracy_sd = sd(nativeAccuracy),
            n = n(),
            rawAccuracy_se = rawAccuracy_sd / sqrt(n), nativeAccuracy_se = nativeAccuracy_sd / sqrt(n),
            .groups = 'keep')

print(categoryAccuracy_tidy)

ggplot(data = categoryAccuracy_tidy, 
       aes(x = as.factor(session), y = nativeAccuracy_mean, fill = group)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme_classic() +
  scale_fill_manual(values=c('grey52','darkorange1'), labels=c('Blind', 'Sighted')) +
  scale_y_continuous(limits=c(0, 0.85), breaks=seq(0, 0.85, 0.1), expand = c(0,0)) +
  scale_x_discrete(limits=c("gorilla", "matlab"),
                   labels = c("Gorilla", "Matlab")) + 
  geom_errorbar(aes(ymin = nativeAccuracy_mean - nativeAccuracy_se,
                    ymax = nativeAccuracy_mean + nativeAccuracy_se),
                width = .08,
                position = position_dodge(.9)) +
  labs(x="Session", y="Accuracy") +
  ggtitle("Native Accuracy") +
  theme(
    text=element_text(size=12), 
    axis.line = element_line(size = 0.6), 
    axis.text.x = element_text(size=10,colour="black",
                               angle = 13,
                               vjust = .5, 
                               hjust = 0.5), 
    axis.text.y = element_text(size=10, colour='black'), 
    legend.title=element_blank())+
  geom_hline(yintercept=c(0.5), linetype="dashed", colour="black", size=0.3)

```

*We input main effect of experimental modalitiy and category,and the interaction between the two. We input as well random intercept by subject, group, item (each stimulus) and language (each language of the foreign words)

### Native Accuracy

```{r}

nativeResult <- glmer(nativeAccuracy ~ session + category + category*session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus), 
      data = allData,
      family = binomial,
      control = glmerControl(optimizer="bobyqa"))

summary(nativeResult)

Anova(nativeResult)

# calculate odd ratio and confidence intervals
se <- sqrt(diag(vcov(nativeResult)))
tab <- cbind(Est = fixef(nativeResult),
             LL = fixef(nativeResult) - 1.96 * se,
             UL = fixef(nativeResult) + 1.96 * se)

# print odd-ratio and confidence interval
print(exp(tab[2:4,]), digits=3)

```

## Post hoc analyses in case we see a difference between groups - Native accuracy

### Category 1: size

```{r}

cat1 = glmer(nativeAccuracy ~ session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus),
      subset(allData, category == 1),
      family = binomial,
      control=glmerControl(optimizer="bobyqa"))

summary(cat1)

Anova(cat1)

catOne_gorilla <- subset(allData, category == 1 & session == 'gorilla')
binom.test(sum(catOne_gorilla$grade), nrow(catOne_gorilla), p = .5)

catOne_matlab <- subset(allData, category == 1 & session == 'matlab')
binom.test(sum(catOne_matlab$grade), nrow(catOne_matlab), p = .5)

```

### Category 3: brightness

```{r}

cat3 = glmer(nativeAccuracy ~ session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus),
      subset(allData, category == 3),
      family = binomial,
      control=glmerControl(optimizer="bobyqa"))

summary(cat3)

Anova(cat3)

catThree_gorilla <- subset(allData, category == 3 & session == 'gorilla')
binom.test(sum(catThree_gorilla$grade), nrow(catThree_gorilla), p = .5)

catThree_matlab <- subset(allData, category == 3 & session == 'matlab')
binom.test(sum(catThree_matlab$grade), nrow(catThree_matlab), p = .5)

```

### Category 4: loudness

```{r}

cat4 = glmer(nativeAccuracy ~ session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus),
      subset(allData, category == 4),
      family = binomial,
      control=glmerControl(optimizer="bobyqa"))

summary(cat4)

Anova(cat4)

catFour_gorilla <- subset(allData, category == 4 & session == 'gorilla')
binom.test(sum(catFour_gorilla$grade), nrow(catFour_gorilla), p = .5)

catFour_matlab <- subset(allData, category == 4 & session == 'matlab')
binom.test(sum(catFour_matlab$grade), nrow(catFour_matlab), p = .5)

```

### Category 6: velocity

```{r}

cat6 = glmer(nativeAccuracy ~ session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus),
      subset(allData, category == 6),
      family = binomial,
      control=glmerControl(optimizer="bobyqa"))

summary(cat6)

Anova(cat6)

catSix_gorilla <- subset(allData, category == 6 & session == 'gorilla')
binom.test(sum(catSix_gorilla$grade), nrow(catSix_gorilla), p = .5)

catSix_matlab <- subset(allData, category == 6 & session == 'matlab')
binom.test(sum(catSix_matlab$grade), nrow(catSix_matlab), p = .5)

```

### Category 7: movement

```{r}

cat7 = glmer(nativeAccuracy ~ session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus),
      subset(allData, category == 7),
      family = binomial,
      control=glmerControl(optimizer="bobyqa"))

summary(cat7)

Anova(cat7)

catSeven_gorilla <- subset(allData, category == 7 & session == 'gorilla')
binom.test(sum(catSeven_gorilla$grade), nrow(catSeven_gorilla), p = .5)

catSeven_matlab <- subset(allData, category == 7 & session == 'matlab')
binom.test(sum(catSeven_matlab$grade), nrow(catSeven_matlab), p = .5)

```

### Category 8: space

```{r}

cat8 = glmer(nativeAccuracy ~ session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus),
      subset(allData, category == 8),
      family = binomial,
      control=glmerControl(optimizer="bobyqa"))

summary(cat8)

Anova(cat8)

catEight_gorilla <- subset(allData, category == 8 & session == 'gorilla')
binom.test(sum(catEight_gorilla$grade), nrow(catEight_gorilla), p = .5)

catEight_matlab <- subset(allData, category == 8 & session == 'matlab')
binom.test(sum(catEight_matlab$grade), nrow(catEight_matlab), p = .5)

```

### Category 9: distance

```{r}

cat9 = glmer(nativeAccuracy ~ session + (1 | group) +
        (1 | pseudoid) + (1 | language:stimulus),
      subset(allData, category == 9),
      family = binomial,
      control=glmerControl(optimizer="bobyqa"))

summary(cat9)

Anova(cat9)

catNine_gorilla <- subset(allData, category == 9 & session == 'gorilla')
binom.test(sum(catNine_gorilla$grade), nrow(catNine_gorilla), p = .5)

catNine_matlab <- subset(allData, category == 9 & session == 'matlab')
binom.test(sum(catNine_matlab$grade), nrow(catNine_matlab), p = .5)

```
