---
title: "Early blindness does not impair the detection of sound symbolic associations in natural language"
author: "Marco Barilari, Minah Chang and Roberto Bottini"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
# Load and install (if missing) the recquired packages
list.of.packages <- c("tidyverse", 
                      "knitr", 
                      "reshape2", 
                      "lme4",
                      "nlme",
                      "car")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(list.of.packages, library, character.only = TRUE)

# set script folder as working dir
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# set to true to show the code chunk
knitr::opts_chunk$set(echo = TRUE)
```

# Data Ananlysys for the paper xxx (under review)

## Data preparation

* All Distractor trials were removed (4186 values) and not considered for further analyses.

### Load processed and merged data

* 1. subID - subject id 
* 2. group - 'EB' either early or congenital blind, 'SC' sighted controls
* 3. session - it refers to the experiment version , either 'matlab' or 'gorilla'
* 4. category [integers 1, 3, 4, 6, 7, 8, 9]
    + category 1: size
    + category 3: brightness
    + category 4: loudness
    + category 6: velocity
    + category 7: movement
    + category 8: space
    + category 9: distance
* 5. wordcode - it refers to the identifier code of each stimulus, the code structure is speaker's gender + lamguage + xxx 
* 6. language - language of origin of the stimulus
* 7. stimulus - sound transcription of the stimulus
* 8. meaning - meaning of the sitmulus refered the general synonim
* 9. value -  'good' or 'good_op',see note below.
* 10. response - participants response as generla synonim 
* 11. grade - response coded as correct (1) or incorrect (0) compared to the 'meaning' column

*Note about 'value' header (native accuracy): given that some words in specific native languages groups can "sound" with the opposite meaning as presented in Tzeng et al., (2017), we recompute the accuracy for those stimuli that in our prestudy were mapped systematically to the opposite meaning. For these subset of stimuli, we consider the response correct if it does not match the original meaning and incorrect if it does. The 9th column "value" indicates whether the accurcy should be calculated as the word's meaning (good), or the opposite meaning (goodop). Raw accuracy rate is calculated without taking into account the value of good or goodop.

```{r}

allData<-read.csv("allData.csv", header = TRUE)

```

### Check that all the participants have the same number of responses per category

```{r}

allData %>% 
  group_by(subID, group, category) %>% 
  count() %>% 
  pivot_wider(names_from = category, values_from = n) %>% 
  print(n = 46)

```

## Descriptive analysis

### Compute the over all native and raw accuracy

```{r}

allData$rawAccuracy <- as.integer(allData$response == allData$meaning)

allData$nativeAccuracy <- ifelse(allData$value == 'goodop', 
                                 abs(allData$rawAccuracy - 1), 
                                 allData$rawAccuracy + 0)
allData %>%
  group_by(group) %>%
  summarize(raw_acc = mean(rawAccuracy), 
            native_acc = mean(nativeAccuracy), 
            .groups = 'keep')

```


### Accuracy per subjects and category 

```{r}

categoryAccuracySubj_tidy <- allData %>%
  group_by(subID, group, category) %>%
  summarize(rawAccuracy_mean = mean(rawAccuracy), nativeAccuracy_mean = mean(nativeAccuracy), 
            rawAccuracy_sd = sd(rawAccuracy), nativeAccuracy_sd = sd(nativeAccuracy),
            n = n(),
            rawAccuracy_se = rawAccuracy_sd / sqrt(n), nativeAccuracy_se = nativeAccuracy_sd / sqrt(n),
            .groups = 'keep')

print(categoryAccuracySubj_tidy)

```

### Accuracy per group and category 

```{r}
categoryAccuracyGroup_tidy <- categoryAccuracySubj_tidy %>%
  group_by(group, category) %>%
  summarize(nativeAccuracy_mean = mean(nativeAccuracy_mean),
            nativeAccuracy_sd = sd(nativeAccuracy_mean),
            n = n(),
            nativeAccuracy_se = nativeAccuracy_sd / sqrt(n),
            .groups = 'keep')

print(categoryAccuracyGroup_tidy)

```

### Plot of the native accuracy per category.

```{r}

ggplot() +
  geom_jitter(data = categoryAccuracySubj_tidy, 
              aes(x = as.factor(category), 
                  y = nativeAccuracy_mean, 
                  color = group),
              position = position_jitterdodge(),
              shape = 1,
              size = 2.5,
              stroke = 1,
              alpha = .5,
              show.legend = F) +
  geom_errorbar(data = categoryAccuracyGroup_tidy,
                aes(y = nativeAccuracy_mean,
                    x = as.factor(category),
                    ymin = nativeAccuracy_mean - nativeAccuracy_se,
                    ymax = nativeAccuracy_mean + nativeAccuracy_se,
                    color = group),
                width = .08,
                alpha = .6,
                position = position_jitterdodge(0),
                show.legend = F) +
  geom_point(data = categoryAccuracyGroup_tidy,
              aes(y = nativeAccuracy_mean,
                  x = as.factor(category),
                  color = group),
              position = position_jitterdodge(0),
              shape = 18,
              size = 6,
              alpha = .8) +
  theme_classic() +
  scale_color_manual(values=c('darkblue','darkorange1'), labels=c('Blind', 'Sighted')) +
  scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, 0.10), expand = c(0,0)) +
  scale_x_discrete(limits=c("1", "3", "4", "6", "7", "8", "9"),
                   labels = c("size", "brightness", "loudness", "velocity", "movement", "space", "distance")) + 
  labs(x="Category", y="Accuracy") +
  ggtitle("Native Accuracy") +
  theme(
    text=element_text(size=14), 
    axis.line = element_line(size = 0.6), 
    axis.text.x = element_text(size=12,colour="black",
                               angle = 13,
                               vjust = .5, 
                               hjust = 0.5), 
    axis.text.y = element_text(size=12, 
                               colour='black'), 
    legend.title=element_blank())+
  geom_hline(yintercept=c(0.5), linetype="dashed", colour="black", size=0.3) 

# ggsave('categories_accuracy.tiff', 
#        device="tiff", 
#        units="in",
#        width=9.54,
#        height=5.54, 
#        dpi=300) 
```

## Statistical analyses

### Prepare the data

```{r}

summary(allData)
allData$subID <- factor(allData$subID)
allData$category <- factor(allData$category)
allData$group <- factor(allData$group)
allData$session <- factor(allData$session)
allData$language <- factor(allData$language)
allData$stimulus <- factor(allData$stimulus)

```

### Experiment modality check (Matlab vs Gorilla)

```{r}
sessionResults <- glmer(nativeAccuracy ~ session +
        (1 | subID) + (1 | language:stimulus), 
      data = allData,
      family = binomial,
      control = glmerControl(optimizer="bobyqa"))

summary(sessionResults)

Anova(sessionResults)
  
```

### Category by Group interaction

Are blind and sighted controls better than chance at detemining the meaning of the presented foreing words? Is there any difference between the two groups?

*We input main effect of semantic category and group and the interaction bewteen the two. We input as well random intercept by subject, item (each stimulus) and language (each language of the foreign words)

```{r}

nativeResults <- glmer(nativeAccuracy ~ session + category + group + session*category*group +
        (1 + category | subID) + (1 + category | language:stimulus), 
      data = allData,
      family = binomial,
      control = glmerControl(optimizer="bobyqa"))

summary(nativeResults)

Anova(nativeResults)

# calculate odd ratio and confidence intervals
se <- sqrt(diag(vcov(nativeResults)))
tab <- cbind(Est = fixef(nativeResults),
             LL = fixef(nativeResults) - 1.96 * se,
             UL = fixef(nativeResults) + 1.96 * se)

# print odd-ratio and confidence interval
print(exp(tab[]), digits=3)

```

### Groups accuracy agaisnt chance

```{r}

native_EB <- subset(allData, group == 'EB')
binom.test(sum(native_EB$grade), nrow(native_EB), p = .5)

native_SC <- subset(allData, group == 'SC')
binom.test(sum(native_SC$grade), nrow(native_SC), p = .5)
```


## Post hoc analyses

### Category 1: size

```{r}

catOne_EB <- subset(allData, category == 1 & group == 'EB')
binom.test(sum(catOne_EB$grade), nrow(catOne_EB), p = .5)

catOne_SC <- subset(allData, category == 1 & group == 'SC')
binom.test(sum(catOne_SC$grade), nrow(catOne_SC), p = .5)

```


### Category 3: brightness

```{r}

catThree_EB <- subset(allData, category == 3 & group == 'EB')
binom.test(sum(catThree_EB$grade), nrow(catThree_EB), p = .5)

catThree_SC <- subset(allData, category == 3 & group == 'SC')
binom.test(sum(catThree_SC$grade), nrow(catThree_SC), p = .5)

```


### Category 4: loudness

```{r}

catFour_EB <- subset(allData, category == 4 & group == 'EB')
binom.test(sum(catFour_EB$grade), nrow(catFour_EB), p = .5)

catFour_SC <- subset(allData, category == 4 & group == 'SC')
binom.test(sum(catFour_SC$grade), nrow(catFour_SC), p = .5)

```


### Category 6: velocity

```{r}

catSix_EB <- subset(allData, category == 6 & group == 'EB')
binom.test(sum(catSix_EB$grade), nrow(catSix_EB), p = .5)

catSix_SC <- subset(allData, category == 6 & group == 'SC')
binom.test(sum(catSix_SC$grade), nrow(catSix_SC), p = .5)

```


### Category 7: movement

```{r}

catSeven_EB <- subset(allData, category == 7 & group == 'EB')
binom.test(sum(catSeven_EB$grade), nrow(catSeven_EB), p = .5)

catSeven_SC <- subset(allData, category == 7 & group == 'SC')
binom.test(sum(catSeven_SC$grade), nrow(catSeven_SC), p = .5)

```


### Category 8: space

```{r}

catEight_EB <- subset(allData, category == 8 & group == 'EB')
binom.test(sum(catEight_EB$grade), nrow(catEight_EB), p = .5)

catEight_SC <- subset(allData, category == 8 & group == 'SC')
binom.test(sum(catEight_SC$grade), nrow(catEight_SC), p = .5)

```


### Category 9: distance

```{r}

catNine_EB <- subset(allData, category == 9 & group == 'EB')
binom.test(sum(catNine_EB$grade), nrow(catNine_EB), p = .5)

catNine_SC <- subset(allData, category == 9 & group == 'SC')
binom.test(sum(catNine_SC$grade), nrow(catNine_SC), p = .5)

```



