---
title: "Guess Meaning analysis"
author: "marcobarilari"
date: "01/23/2020"
output: html_document
---

```{r setup, include=FALSE}
# Load and install, if missing, the recquired packages
list.of.packages <- c("tidyverse")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(list.of.packages, library, character.only = TRUE)

knitr::opts_chunk$set(echo = TRUE)
```

## Load and reshape the data

Load the data and match the columns, then bind them in one dataframe

```{r}

path_raw <- '/Users/barilari/Desktop/GuessMeaning_final-data/raw'

raw_allGorilla <- read.table(paste(path_raw, 'raw_allGorilla.tsv', sep = '/'), 
                                 header = T, 
                                 sep = '\t')

raw_allMatlab <- read.table(paste(path_raw, 'raw_allMatlab.tsv', sep = '/'), 
                                header = T, 
                                sep = '\t')

allGorilla <- raw_allGorilla %>% 
  select(-rotation)

allMatlab <- raw_allMatlab %>% 
  select(-subjNb, -expTrialNb, -buttonOrder, -categoryTrialNb)
```

Delete unmatched subjects

```{r}
allGorilla <- allGorilla %>% 
  filter(!pseudoID %in% c('19871012MIFR', '19700823EICS', '19880227DNTM', '19970409MRLM'))

allMatlab <- allMatlab %>% 
  filter(!pseudoID %in% c('19890803ABBN', '19930320GOSC', '19780909ANMN', '19920215AGMC'))
```

Exclude distractors

```{r}
allGorilla <- allGorilla %>% 
  filter(value != 'distractor') %>% 
  droplevels()

allMatlab <- allMatlab %>% 
  filter(value != 'distractor') %>% 
  droplevels()
```

Merge all together

```{r}
allData <- bind_rows(allGorilla, allMatlab)
```

## Look the at the data

Summary of the data 

```{r}
summary(allData)
```

Check that all the participants have the same number of responses per category

```{r}
allData %>% 
  group_by(pseudoID, group, category) %>% 
  count() %>% 
  pivot_wider(names_from = category, values_from = n) %>% 
  print(n = 46)
```


## Comparison between group

###Over all accuracy

Compute the "raw" accuracy 

```{r}
allData$rawAccuracy <- as.integer(allData$response == allData$meaning)

allData %>%
  group_by(group) %>%
  summarize(rawAccuracy = mean(rawAccuracy), 
            .groups = 'keep')
```

Compute the accuracy given that some word in specific native language groups can "sound" with the opposite meaning

from Tzeng et al 2017 Cognitive Science:

>###5.2. Systematic versus “correct” sound-to-meaning mappings
Of the 80 foreign sound symbolic words, 35 were reverse mappings such that listeners, as a group, reliably mapped them to the opposite of the words’ meanings. That listeners consistently mapped the words to the “wrong” meaning within the dimension for a significant proportion of our items which raises the question of what the functional significance of such sound-to-meaning mappings might be. Indeed, there appears to be some probability that overreliance on sound-to-meaning cues could lead listeners to an incorrect interpretation of specific word meaning (e.g., big vs. small). It may be that the functional utility of sound symbolic cues to meaning is to help constrain the meaning to a particular dimension but does not necessarily determine the specific meaning within this dimension or semantic domain.

Therefor, from our prestudy, we determined wihich stimuli were mapped systematically to the opposite meaning in an indipented sample of native italian speakers. Below we re-compute the accuracy for those stimuli (if the response does not match the meaning we code it as a correct reposnse and vice versa)

```{r}
allData$nativeAccuracy <- ifelse(allData$value == 'good_op', 
                                 abs(allData$rawAccuracy - 1), 
                                 allData$rawAccuracy + 0)

allData %>%
  group_by(group) %>%
  summarize(native_acc = mean(nativeAccuracy), 
            .groups = 'keep')
```

Display them together

```{r}
allData %>%
  group_by(group) %>%
  summarize(raw_acc = mean(rawAccuracy), 
            native_acc = mean(nativeAccuracy), 
            .groups = 'keep')
```


###Category accuracy per experiment version

as a reminder:

- categroy 1: size
- categroy 3: brightness
- categroy 4: loudness
- categroy 6: velocity
- categroy 7: movement
- categroy 8: space
- categroy 9: distance

```{r}
categoryAccuracy_wide <- allData %>%
  group_by(category, group) %>%
  summarize(rawAccuracy = mean(rawAccuracy), nativeAccuracy = mean(nativeAccuracy), 
            .groups = 'keep') %>% 
  pivot_wider(names_from = group, values_from = c(rawAccuracy, nativeAccuracy))

print(categoryAccuracy_wide)

```

Plot the accuracies per category and experiment version


```{r}

categoryAccuracy_tidy <- allData %>%
  group_by(group, category) %>%
  summarize(rawAccuracy = mean(rawAccuracy), nativeAccuracy = mean(nativeAccuracy), 
            .groups = 'keep')

ggplot(data = categoryAccuracy_tidy, 
       aes(x = as.factor(category), y = rawAccuracy, fill = group)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme_classic() +
  scale_fill_manual(values=c('red','blue'), labels=c('EB', 'SC')) +
  scale_y_continuous(limits=c(0, 0.85), breaks=seq(0, 0.85, 0.1), expand = c(0,0)) +
  scale_x_discrete(limits=c("1", "3", "4", "6", "7", "8", "9"),
                   labels = c("size", "brightness", "loudness", "velocity", "movement", "space", "distance")) + 
  labs(x="Semantic categories", y="Accuracy") +
  ggtitle("Raw accuracy") +
  theme(
    text=element_text(size=18), 
    axis.line = element_line(size = 0.6), 
    axis.text.x = element_text(size=14,colour="black",
                               angle = 15,
                               vjust = .5, 
                               hjust = 0.5),
    axis.text.y = element_text(size=16, colour='black'), 
    legend.title=element_blank())+
  geom_hline(yintercept=c(0.5), linetype="dashed", colour="black", size=0.3)

ggplot(data = categoryAccuracy_tidy, 
       aes(x = as.factor(category), y = nativeAccuracy, fill = group)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme_classic() +
  scale_fill_manual(values=c('red','blue'), labels=c('EB', 'SC')) +
  scale_y_continuous(limits=c(0, 0.85), breaks=seq(0, 0.85, 0.1), expand = c(0,0)) +
  scale_x_discrete(limits=c("1", "3", "4", "6", "7", "8", "9"),
                   labels = c("size", "brightness", "loudness", "velocity", "movement", "space", "distance")) + 
  labs(x="Semantic categories", y="Accuracy") +
  ggtitle("Native accuracy") +
  theme(
    text=element_text(size=18), 
    axis.line = element_line(size = 0.6), 
    axis.text.x = element_text(size=14,colour="black",
                               angle = 15,
                               vjust = .5, 
                               hjust = 0.5), 
    axis.text.y = element_text(size=16, colour='black'), 
    legend.title=element_blank())+
  geom_hline(yintercept=c(0.5), linetype="dashed", colour="black", size=0.3)

```



